---
title: Scatterplot Colors
output: 
  html_document:
    css: color_style.css
    toc: true
    toc_float: true
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

library(knitr)
library(tidyverse)
library(plotly)

df <- readRDS("data/cvl_data.RDS")
```

Michael pointed out that the scatterplot colors had some overlap with the bivariate color palette that could be confusing -- folks might make unintended connections between the two palettes. I spent some time looking through some other qualitative palette options (using [Emil Hvitfeldt's fantabulous curated list](https://github.com/EmilHvitfeldt/r-color-palettes) of R color palettes as a guide), but I need other eyes on the possibilities. I've added the bivariate legend on the sidebar so you can see the scatterplot palette in relation to all of the options.

## RColorBrewer::Dark2
This is the current palette; note how the off-diagonal colors on the bivarate color matrix have close analogues in this Greene, Fluvanna, and Albemarle colors.

```{r dark2}
library(RColorBrewer)
pal <- brewer.pal(6, "Dark2")

plot_ly(data = df, 
                     x = ~tree_can, 
                     y = ~whiteE, 
                     type = 'scatter',
                     mode = 'markers', # to remove mode warning
                     fill = ~'', # to remove line.width error
                     size = ~totalpopE, 
                     sizes = c(1, 500),
                     color = ~countyname, 
                     colors = pal, 
                     alpha = .75,
                     text = paste0("County: ", df$countyname, "<br>",
                                   "Population: ", df$totalpopE, "<br>",
                                   "Census tract: ", df$tract, "<br>",
                                   "X: ", df$tree_can, "<br>",
                                   "Y: ", df$whiteE, "<br>"),
                     hoverinfo = "text") %>% 
  layout(xaxis = list(title = "Tree Canopy", showticklabels = TRUE),
         yaxis = list(title = "Percent White", showtickLabels = TRUE),
         legend = list(orientation = "h", x = 0, y = -0.2))

```

## RColorBrewer::Set1
An alternative color palette: I like this one and think it overlaps less, but it feels a tad too cheery to me.
```{r set1}
pal <- brewer.pal(6, "Set1")

plot_ly(data = df, 
                     x = ~tree_can, 
                     y = ~whiteE, 
                     type = 'scatter',
                     mode = 'markers', # to remove mode warning
                     fill = ~'', # to remove line.width error
                     size = ~totalpopE, 
                     sizes = c(1, 500),
                     color = ~countyname, 
                     colors = pal, 
                     alpha = .75,
                     text = paste0("County: ", df$countyname, "<br>",
                                   "Population: ", df$totalpopE, "<br>",
                                   "Census tract: ", df$tract, "<br>",
                                   "X: ", df$tree_can, "<br>",
                                   "Y: ", df$whiteE, "<br>"),
                     hoverinfo = "text") %>% 
  layout(xaxis = list(title = "Tree Canopy", showticklabels = TRUE),
         yaxis = list(title = "Percent White", showtickLabels = TRUE),
         legend = list(orientation = "h", x = 0, y = -0.2))

```

## jcolors::pal5
The color assigned to Charlottesville in this one may be a bit light.
```{r pal5}
library(jcolors)
pal <- unname(jcolors("pal5")[1:6])

plot_ly(data = df, 
                     x = ~tree_can, 
                     y = ~whiteE, 
                     type = 'scatter',
                     mode = 'markers', # to remove mode warning
                     fill = ~'', # to remove line.width error
                     size = ~totalpopE, 
                     sizes = c(1, 500),
                     color = ~countyname, 
                     colors = pal, 
                     alpha = .75,
                     text = paste0("County: ", df$countyname, "<br>",
                                   "Population: ", df$totalpopE, "<br>",
                                   "Census tract: ", df$tract, "<br>",
                                   "X: ", df$tree_can, "<br>",
                                   "Y: ", df$whiteE, "<br>"),
                     hoverinfo = "text") %>% 
  layout(xaxis = list(title = "Tree Canopy", showticklabels = TRUE),
         yaxis = list(title = "Percent White", showtickLabels = TRUE),
         legend = list(orientation = "h", x = 0, y = -0.2))
```

## jcolors::rainbow
I selected 6 of the 9 colors from this palette that seemed less overlappy with the bivariate legend, and assigned them out of order (so Greene would be green).
```{r rainbow}
pal <- unname(jcolors("rainbow")[c(8,2,6,4,1,3)])

plot_ly(data = df, 
                     x = ~tree_can, 
                     y = ~whiteE, 
                     type = 'scatter',
                     mode = 'markers', # to remove mode warning
                     fill = ~'', # to remove line.width error
                     size = ~totalpopE, 
                     sizes = c(1, 500),
                     color = ~countyname, 
                     colors = pal, 
                     alpha = .75,
                     text = paste0("County: ", df$countyname, "<br>",
                                   "Population: ", df$totalpopE, "<br>",
                                   "Census tract: ", df$tract, "<br>",
                                   "X: ", df$tree_can, "<br>",
                                   "Y: ", df$whiteE, "<br>"),
                     hoverinfo = "text") %>% 
  layout(xaxis = list(title = "Tree Canopy", showticklabels = TRUE),
         yaxis = list(title = "Percent White", showtickLabels = TRUE),
         legend = list(orientation = "h", x = 0, y = -0.2))
```

## ggthemes::few
Chose six of eight colors from this palette (excluding pink and purple).
```{r few}
library(ggthemes)
pal <- c("#265dab", "#df5c24", "#c7b42e", "#059748",
         "#cb2027", "#9d722a")

plot_ly(data = df, 
                     x = ~tree_can, 
                     y = ~whiteE, 
                     type = 'scatter',
                     mode = 'markers', # to remove mode warning
                     fill = ~'', # to remove line.width error
                     size = ~totalpopE, 
                     sizes = c(1, 500),
                     color = ~countyname, 
                     colors = pal, 
                     alpha = .75,
                     text = paste0("County: ", df$countyname, "<br>",
                                   "Population: ", df$totalpopE, "<br>",
                                   "Census tract: ", df$tract, "<br>",
                                   "X: ", df$tree_can, "<br>",
                                   "Y: ", df$whiteE, "<br>"),
                     hoverinfo = "text") %>% 
  layout(xaxis = list(title = "Tree Canopy", showticklabels = TRUE),
         yaxis = list(title = "Percent White", showtickLabels = TRUE),
         legend = list(orientation = "h", x = 0, y = -0.2))
```


## ggthemes::calc
```{r calc}
pal <- c("#004586", "#ff420e", "#ffd320", "#579d1c",
         "#7e0021", "#83caff")

plot_ly(data = df, 
                     x = ~tree_can, 
                     y = ~whiteE, 
                     type = 'scatter',
                     mode = 'markers', # to remove mode warning
                     fill = ~'', # to remove line.width error
                     size = ~totalpopE, 
                     sizes = c(1, 500),
                     color = ~countyname, 
                     colors = pal, 
                     alpha = .75,
                     text = paste0("County: ", df$countyname, "<br>",
                                   "Population: ", df$totalpopE, "<br>",
                                   "Census tract: ", df$tract, "<br>",
                                   "X: ", df$tree_can, "<br>",
                                   "Y: ", df$whiteE, "<br>"),
                     hoverinfo = "text") %>% 
  layout(xaxis = list(title = "Tree Canopy", showticklabels = TRUE),
         yaxis = list(title = "Percent White", showtickLabels = TRUE),
         legend = list(orientation = "h", x = 0, y = -0.2))
```

## ggsci::startrek
I quite like this one, and not just because I'm very excited about Star Trek.
```{r startrek}
library(ggsci)
pal <- c("#7c878eff", "#ffcd00ff", "#00b5e2ff", "#00af66ff",
         "#5c88daff", "#84bd00ff")

plot_ly(data = df, 
                     x = ~tree_can, 
                     y = ~whiteE, 
                     type = 'scatter',
                     mode = 'markers', # to remove mode warning
                     fill = ~'', # to remove line.width error
                     size = ~totalpopE, 
                     sizes = c(1, 500),
                     color = ~countyname, 
                     colors = pal, 
                     alpha = .75,
                     text = paste0("County: ", df$countyname, "<br>",
                                   "Population: ", df$totalpopE, "<br>",
                                   "Census tract: ", df$tract, "<br>",
                                   "X: ", df$tree_can, "<br>",
                                   "Y: ", df$whiteE, "<br>"),
                     hoverinfo = "text") %>% 
  layout(xaxis = list(title = "Tree Canopy", showticklabels = TRUE),
         yaxis = list(title = "Percent White", showtickLabels = TRUE),
         legend = list(orientation = "h", x = 0, y = -0.2))
```


## ghibli::Mononoke
Now for a series from the ghibli package, with palettes I've long wanted to use, but have yet to find the right project for... could this be the one?

The Ghibil Fandom wiki describes Princess Mononoke as a story about "the struggle between the supernatural guardians of a forest and the humans who consume its resources."
```{r mononoke}
library(ghibli)
pal <- ghibli_palette("MononokeMedium")[2:7]

plot_ly(data = df, 
                     x = ~tree_can, 
                     y = ~whiteE, 
                     type = 'scatter',
                     mode = 'markers', # to remove mode warning
                     fill = ~'', # to remove line.width error
                     size = ~totalpopE, 
                     sizes = c(1, 500),
                     color = ~countyname, 
                     colors = pal, 
                     alpha = .75,
                     text = paste0("County: ", df$countyname, "<br>",
                                   "Population: ", df$totalpopE, "<br>",
                                   "Census tract: ", df$tract, "<br>",
                                   "X: ", df$tree_can, "<br>",
                                   "Y: ", df$whiteE, "<br>"),
                     hoverinfo = "text") %>% 
  layout(xaxis = list(title = "Tree Canopy", showticklabels = TRUE),
         yaxis = list(title = "Percent White", showtickLabels = TRUE),
         legend = list(orientation = "h", x = 0, y = -0.2))
```

## ghibli::Kiki
Kiki's Delivery Service is not as thematically aligned, but pretty.
```{r kiki}
pal <- ghibli_palette("KikiMedium")[2:7]

plot_ly(data = df, 
                     x = ~tree_can, 
                     y = ~whiteE, 
                     type = 'scatter',
                     mode = 'markers', # to remove mode warning
                     fill = ~'', # to remove line.width error
                     size = ~totalpopE, 
                     sizes = c(1, 500),
                     color = ~countyname, 
                     colors = pal, 
                     alpha = .75,
                     text = paste0("County: ", df$countyname, "<br>",
                                   "Population: ", df$totalpopE, "<br>",
                                   "Census tract: ", df$tract, "<br>",
                                   "X: ", df$tree_can, "<br>",
                                   "Y: ", df$whiteE, "<br>"),
                     hoverinfo = "text") %>% 
  layout(xaxis = list(title = "Tree Canopy", showticklabels = TRUE),
         yaxis = list(title = "Percent White", showtickLabels = TRUE),
         legend = list(orientation = "h", x = 0, y = -0.2))
```

## ghibli::Spirited
Spirited Away, a young girl navigates the spirit world and saves her parents; I'm actually like these muted palettes, but wondering if they are visually distinct enough (though perhaps strong visual distinctions between the localities isn't necessary).
```{r spirited}
pal <- ghibli_palette("SpiritedMedium")[2:7]

plot_ly(data = df, 
                     x = ~tree_can, 
                     y = ~whiteE, 
                     type = 'scatter',
                     mode = 'markers', # to remove mode warning
                     fill = ~'', # to remove line.width error
                     size = ~totalpopE, 
                     sizes = c(1, 500),
                     color = ~countyname, 
                     colors = pal, 
                     alpha = .75,
                     text = paste0("County: ", df$countyname, "<br>",
                                   "Population: ", df$totalpopE, "<br>",
                                   "Census tract: ", df$tract, "<br>",
                                   "X: ", df$tree_can, "<br>",
                                   "Y: ", df$whiteE, "<br>"),
                     hoverinfo = "text") %>% 
  layout(xaxis = list(title = "Tree Canopy", showticklabels = TRUE),
         yaxis = list(title = "Percent White", showtickLabels = TRUE),
         legend = list(orientation = "h", x = 0, y = -0.2))
```

## ghibli::Ponyo
A fish becomes human and the balance of nature is temporarily disrupted
```{r ponyo}
pal <- ghibli_palette("PonyoMedium")[2:7]

plot_ly(data = df, 
                     x = ~tree_can, 
                     y = ~whiteE, 
                     type = 'scatter',
                     mode = 'markers', # to remove mode warning
                     fill = ~'', # to remove line.width error
                     size = ~totalpopE, 
                     sizes = c(1, 500),
                     color = ~countyname, 
                     colors = pal, 
                     alpha = .75,
                     text = paste0("County: ", df$countyname, "<br>",
                                   "Population: ", df$totalpopE, "<br>",
                                   "Census tract: ", df$tract, "<br>",
                                   "X: ", df$tree_can, "<br>",
                                   "Y: ", df$whiteE, "<br>"),
                     hoverinfo = "text") %>% 
  layout(xaxis = list(title = "Tree Canopy", showticklabels = TRUE),
         yaxis = list(title = "Percent White", showtickLabels = TRUE),
         legend = list(orientation = "h", x = 0, y = -0.2))
```

## ghibli::Totoro
My Neighbor Totoro, wood spirits and environmentalism - I love this movie so much.
```{r totoro}
pal <- ghibli_palette("TotoroMedium")[2:7]

plot_ly(data = df, 
                     x = ~tree_can, 
                     y = ~whiteE, 
                     type = 'scatter',
                     mode = 'markers', # to remove mode warning
                     fill = ~'', # to remove line.width error
                     size = ~totalpopE, 
                     sizes = c(1, 500),
                     color = ~countyname, 
                     colors = pal, 
                     alpha = .75,
                     text = paste0("County: ", df$countyname, "<br>",
                                   "Population: ", df$totalpopE, "<br>",
                                   "Census tract: ", df$tract, "<br>",
                                   "X: ", df$tree_can, "<br>",
                                   "Y: ", df$whiteE, "<br>"),
                     hoverinfo = "text") %>% 
  layout(xaxis = list(title = "Tree Canopy", showticklabels = TRUE),
         yaxis = list(title = "Percent White", showtickLabels = TRUE),
         legend = list(orientation = "h", x = 0, y = -0.2))
```